CF Registry Analysis

### 2023-12-07 Christine Chan/CF Registry/cf_registry_analysis.qmd
registry survival analysis data formatting

```{r}
#| include: false
library(Hmisc)
library(tidyverse)
library(forestmodel)
library(childsds)
library(gt)
library(gtsummary)
library(forcats)
library(labelled)
library(parallel)
```

```{r}
# Import data and codebook
annual <- read.csv("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
data_dict <- readxl::read_excel("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
# Remove those with T1D or T2D
annual <- annual %>% arrange(eDWID, ReviewYear)
annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
  levels = 1:3,
  labels = c("CFRD-", "CFRD-", "CFRD+")
)
annual$age_2018 <- 2018 - (annual$ReviewYear - annual$Age_YrEnd)
annual <- annual %>%
  group_by(eDWID) %>%
  mutate(age_2018 = as.numeric(names(sort(table(age_2018),
    decreasing = TRUE
  )[1])))
# Get mutations
demogs <- read.csv("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
demogs <- demogs %>% select(eDWID, Gender, MutClass)
annual <- left_join(annual, demogs, by = join_by(eDWID))
# Variable formatting
# If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
# on any supplemental feeding
# Include all kinds of tube feeds
annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
annual$tube_feeding <-
  apply(annual[, tube_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Liver disease
liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
  apply(annual[, liver_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Calculate percentiles
annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$age_perc <- ifelse(annual$Age_YrEnd < 20, annual$Age_YrEnd, 20)
annual$height_perc <- sds(annual$A_height,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "height2_20", ref = cdc.ref, type = "perc"
) * 100
annual$bmi_perc <- sds(annual$bmi,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "bmi", ref = cdc.ref, type = "perc"
) * 100
# Format columns
annual$MutClass <- factor(annual$MutClass,
  levels = c("1-3", "4-5", "Oth", "N/A"),
  labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
  labels = c("0", "1", "2", ">=3"),
  breaks = c(0, 1, 2, 3, Inf), right = F
)
annual$A_FEV1_pct_predicted <- cut(annual$A_FEV1_pct_predicted,
  breaks = c(0, 40, 80, 100, Inf), right = F
)
annual$A_FEV1_pct_predicted <- relevel(annual$A_FEV1_pct_predicted,
  ref = "[100,Inf)"
)
annual$height_perc <- cut(annual$height_perc, breaks = c(0, 10, 25, 50, Inf))
annual$height_perc <- relevel(annual$height_perc, ref = "(50,Inf]")
annual$bmi_perc <- cut(annual$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
annual$bmi_perc <- relevel(annual$bmi_perc, ref = "(50,85]")
annual$age_group <- cut(annual$Age_YrEnd,
  breaks = c(-Inf, 10, 18, 30, Inf),
  right = F
)
annual$age_group <- relevel(annual$age_group, ref = "[10,18)")
# Bugs
bugs <- c(
  "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_pulmonarycomplications1", "A_mycocultureresults3"
)
# Yes/No variables
yn_vars <- c(
  "A_IsOnEnzymes", "A_corticosteroids1", "A_feeding2",
  "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"
)
annual[, yn_vars] <- lapply(annual[, yn_vars], factor,
  levels = 0:1,
  labels = c("No", "Yes")
)
annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
```

@@

@@

## By year

```{r results='asis'}
# Find the year that each person was diagnosed, make wide
first_cfrd <- annual %>%
  filter(cfrd_status_annual == "CFRD+") %>%
  group_by(eDWID) %>%
  slice_head(n = 1) %>%
  select(eDWID, ReviewYear) %>%
  rename(first_cfrd = ReviewYear)
# Make a dataframe for who was at risk each year
at_risk <- expand.grid(eDWID = unique(annual$eDWID), ReviewYear = 2003:2018)
at_risk <- at_risk %>% arrange(eDWID)
at_risk <- at_risk %>%
  full_join(., annual %>%
    select(eDWID, ReviewYear, age_2018, A_hgba1c, cfrd_status_annual),
  by = join_by(eDWID, ReviewYear)
  )
# Combine
inc <- left_join(at_risk, first_cfrd, by = join_by(eDWID))
# Calculate all missing ages
inc <- inc %>%
  group_by(eDWID) %>%
  mutate(age = first(na.omit(age_2018)) - (2018 - ReviewYear)) %>%
  # No missing age in 2018 field, so this indicates whether they were at risk
  # (i.e. seen that year)
  mutate(age_2018 = !is.na(age_2018)) %>%
  rename(at_risk = age_2018) %>%
  ungroup()
# Age group
inc$age_group <- cut(inc$age, c(-Inf, 10, 19, 31, Inf), right = F)
# Remove people from at risk after diagnosis
inc$at_risk[!is.na(inc$first_cfrd) & inc$ReviewYear > inc$first_cfrd] <- FALSE
# Use gtsummary to make a list of tables by year
invisible(lapply(unique(inc$ReviewYear), function(y) {
  d <- inc %>%
    filter(ReviewYear == y, at_risk) %>%
    # Make missing values explicit for the tables
    mutate(
      a1c = factor(!is.na(A_hgba1c), levels = c(F, T), labels = c("No", "Yes")),
      cfrd_status_annual = fct_explicit_na(cfrd_status_annual),
    ) %>%
    select(age_group, cfrd_status_annual, a1c, A_hgba1c) %>%
    tbl_summary(
      by = age_group, missing_text = "(Missing)",
      label = list(
        cfrd_status_annual ~ "Annualized CFRD Status",
        A_hgba1c ~ "Annualized HbA1c", a1c ~ "A1c Collected"
      ),
      statistic = list(all_continuous() ~ "{mean} ({sd})")
    ) %>%
    add_overall()
  cat("\n")
  cat("### ", y)
  cat("\n")
  print(d)
  cat("\n")
}))
```

# Logistic analyses

Any participants with T1D or T2D were excluded from these analyses. Variables that were entirely missing in a given year (e.g., `A_Vx770` and `A_VX809comb` in 2005) were excluded from those models. Participants under age 10 at year end were also excluded due to missing PFT data.

```{r}
# Analysis function
log_reg <- function(year, missingness = F) {
  # Clean data
  df <- annual %>%
    filter(ReviewYear == year, Age_YrEnd >= 10) %>%
    select(
      eDWID, age_group, Gender, bmi_perc, MutClass, A_IsOnEnzymes,
      A_FEV1_pct_predicted, tube_feeding, NumPulmExacerbation,
      A_pseudomonasaeruginosa, A_fungalyeast1, A_mycocultureresults3,
      A_pulmonarycomplications1,
      A_corticosteroids1, cfrd_status_annual, A_Vx770, A_VX809comb,
      liver_disease
    )
  # Assume that missing are CFRD- for these models
  df$cfrd_status_annual[is.na(df$cfrd_status_annual)] <- "CFRD-"
  # Relevel age group since we're excluding kids < 10
  df$age_group <- droplevels(df$age_group)
  # Remove all empty columns
  empty <- which(colSums(is.na(df)) == nrow(df))
  if (length(empty) > 0) {
    df <- df[, -empty]
  }
  # Models
  vars <- colnames(df)[which(!colnames(df) %in%
    c(bugs, liver_fields, "cfrd_status_annual", "eDWID"))]
  base_mod <- as.formula(paste0(
    "cfrd_status_annual~",
    paste0(vars, collapse = "+")
  ))
  # Labels
  labels <- list(
    cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
    Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
    A_FEV1_pct_predicted = "FEV1 percent predicted",
    tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
    A_corticosteroids1 = "Use of corticosteroids",
    NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
    height_perc = "Height percentile", bmi_perc = "BMI percentile",
    A_Vx770 = "Use of CFTR modulator Vx770",
    A_VX809comb = "Use of CFTR modulator Vx809 combination",
    A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
    A_fungalyeast1 = "Aspergillus (any species)",
    A_pulmonarycomplications1 = "ABPA",
    A_mycocultureresults3 = "Microorganisms in myco culture"
  )
  var_label(df[, vars[which(vars %in% colnames(df))]]) <-
    as.character(labels[vars[which(vars %in% colnames(df))]])
  var_label(df[, bugs[which(bugs %in% colnames(df))]]) <-
    as.character(labels[bugs[which(bugs %in% colnames(df))]])
  # Visualize missingness if required
  if (missingness) {
    cat("\n")
    cat("### Missingness")
    cat("\n")
    cat("\n")
    print(naniar::vis_miss(df))
    cat("\n")
  }
  # Simple models
  mod1 <- glm(base_mod, data = df, family = "binomial")
  # Print results
  cat("\n")
  cat("### Base Model")
  cat("\n")
  cat("\n")
  print(forest_model(mod1, exclude_infinite_cis = F))
  cat("\n")
  if ("A_pseudomonasaeruginosa" %in% colnames(df)) {
    mod2 <- glm(update(base_mod, . ~ . + A_pseudomonasaeruginosa),
      data = df,
      family = "binomial"
    )
    cat("\n")
    cat("### P. aeruginosa")
    cat("\n")
    cat("\n")
    print(forest_model(mod2, exclude_infinite_cis = F))
    cat("\n")
  }
  if ("A_fungalyeast1" %in% colnames(df)) {
    mod3 <- glm(update(base_mod, . ~ . + A_fungalyeast1),
      data = df,
      family = "binomial"
    )
    cat("\n")
    cat("### Aspergillus")
    cat("\n")
    cat("\n")
    print(forest_model(mod3, exclude_infinite_cis = F))
    cat("\n")
  }
  if ("A_pulmonarycomplications1" %in% colnames(df)) {
    mod4 <- glm(update(base_mod, . ~ . + A_pulmonarycomplications1),
      data = df,
      family = "binomial"
    )
    cat("\n")
    cat("### ABPA")
    cat("\n")
    cat("\n")
    print(forest_model(mod4, exclude_infinite_cis = F))
    cat("\n")
  }
  if ("A_mycocultureresults3" %in% colnames(df)) {
    mod5 <- glm(update(base_mod, . ~ . + A_mycocultureresults3),
      data = df,
      family = "binomial"
    )
    cat("\n")
    cat("### Mycobacterium")
    cat("\n")
    cat("\n")
    print(forest_model(mod5, exclude_infinite_cis = F))
    cat("\n")
  }
}
```

```{r results='asis',fig.height=10,fig.width=10}
invisible(lapply(unique(annual$ReviewYear), function(y) {
  cat("\n")
  cat("## ", y)
  cat("\n")
  log_reg(y)
  cat("\n")
}))
```

>>>>

```{r}
#| include: false
library(Hmisc)
library(tidyverse)
library(forestmodel)
library(childsds)
library(gt)
library(gtsummary)
library(forcats)
library(labelled)
library(survival)
library(survminer)
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
# All data
encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv",
  na.strings = c("", "N/A")
)
annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
annual <- annual %>% arrange(eDWID, ReviewYear)
# Add age
ages <- annual %>%
  mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd)) %>%
  rename(reviewyear = ReviewYear) %>%
  group_by(eDWID, reviewyear) %>%
  summarise(
    `Age in 2018` = as.numeric(names(sort(table(`Age in 2018`),
      decreasing = TRUE
    )[1])),
    Age = Age_YrEnd[1], .groups = "drop"
  )
encounter <- left_join(encounter, ages, by = c("eDWID", "reviewyear"))
encounter$`Age Group 2018` <- cut(encounter$`Age in 2018`,
  c(-Inf, 10, 19, 31, Inf),
  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
  right = F
)
encounter$`Age Group` <- cut(encounter$Age,
  c(-Inf, 10, 19, 31, Inf),
  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
  right = F
)
# For those with CFRD by registry criteria, find first visit with diagnosis
first_cfrd_cffpr <- encounter %>%
  filter(cfrd_status == 3) %>%
  group_by(eDWID) %>%
  filter(row_number() == 1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_cffpr = reviewyear)
# Add to encounter level data
encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_both, by = c("eDWID", "Age in 2018"))
# Fill down CFRD Yes/No by year
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_cffpr = case_when(
      reviewyear < first_cfrd_cffpr ~ "No",
      reviewyear >= first_cfrd_cffpr ~ "Yes"
    )
  ) %>%
  ungroup()
encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
# Ordered factors
encounter$cfrd_ogtt <- factor(encounter$cfrd_ogtt, levels = c("No", "Yes"), ordered = T)
encounter$cfrd_cffpr <- factor(encounter$cfrd_cffpr, levels = c("No", "Yes"), ordered = T)
encounter$cfrd_both <- factor(encounter$cfrd_both, levels = c("No", "Yes"), ordered = T)
# Remove those missing age
encounter <- encounter %>% filter(!is.na(Age))
```

```{r}
#| label: tbl-incidence_cffpr
#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone"
# For each year, pull everyone with and without CFRD that year
year_inc <- function(df = encounter, year, py = 1000) {
  # Filter and count
  inc <- df %>%
    filter(reviewyear == year, is.na(first_cfrd_cffpr) | first_cfrd_cffpr >= year) %>%
    group_by(eDWID) %>%
    filter(row_number() == 1) %>%
    select(eDWID, `Age Group`, cfrd_cffpr)
  inc <- inc %>%
    group_by(cfrd_cffpr) %>%
    count(`Age Group`, .drop = F)
  inc <- inc[!is.na(inc$`Age Group`), ]
  # Calculations
  new_cases <- sum(inc$n[inc$cfrd_cffpr == "Yes"])
  total_at_risk <- sum(inc$n)
  total_inc <- new_cases / (total_at_risk / py)

  new_cases_10 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "< 10"])
  at_risk_10 <- sum(inc$n[inc$`Age Group` == "< 10"])
  inc_10 <- new_cases_10 / (at_risk_10 / py)

  new_cases_10_18 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "10 - 18"])
  at_risk_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)

  new_cases_19_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "19 - 30"])
  at_risk_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)

  new_cases_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "> 30"])
  at_risk_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
  inc_30 <- new_cases_30 / (at_risk_30 / py)

  # Nice results
  list(
    `Year` = year,
    `Total New Cases` = new_cases,
    `Total At Risk` = total_at_risk,
    `Total Incidence` = total_inc,
    `Total New Cases Age < 10` = new_cases_10,
    `Total At Risk Age < 10` = at_risk_10,
    `Total Incidence Age < 10` = inc_10,
    `Total New Cases Age 10-18` = new_cases_10_18,
    `Total At Risk Age 10-18` = at_risk_10_18,
    `Total Incidence Age 10-18` = inc_10_18,
    `Total New Cases Age 19-30` = new_cases_19_30,
    `Total At Risk Age 19-30` = at_risk_19_30,
    `Total Incidence Age 19-30` = inc_19_30,
    `Total New Cases Age > 30` = new_cases_30,
    `Total At Risk Age > 30` = at_risk_30,
    `Total Incidence Age > 30` = inc_30
  )
}
incidence <- lapply(2003:2018, function(y) {
  year_inc(year = y)
})
incidence <- data.frame(do.call(rbind, incidence), check.names = F)
incidence <- data.frame(lapply(incidence, as.numeric), check.names = F)
kable(incidence, digits = 3)
```

In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases.

```{r}
#| label: tbl-prevalence_cffpr
#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone"
encounter %>%
  group_by(eDWID, reviewyear) %>%
  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
  group_by(reviewyear) %>%
  summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>%
  kable(., digits = 3)
```

```{r}
#| label: tbl-prevalence_cffpr_year
#| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone"
encounter %>%
  group_by(eDWID, reviewyear, `Age Group`) %>%
  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
  group_by(reviewyear, `Age Group`) %>%
  summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>%
  pivot_wider(names_from = `Age Group`, values_from = Prevalence) %>%
  kable(., digits = 3)
```

In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.

# CFRD screening rates

```{r}
#| label: tbl-screening_ogtt
#| tbl-cap: "Yearly OGTT Screening Rates"
year_screen <- function(df = encounter, year) {
  # Filter and count
  inc <- df %>%
    filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
    mutate(ogtt = !is.na(ogttfast_bloodglucose) | !is.na(twohour_bloodglucose)) %>%
    group_by(eDWID) %>%
    summarise(ogtt = max(ogtt), `Age Group` = `Age Group`[1])
  inc <- inc %>%
    group_by(ogtt) %>%
    count(`Age Group`) %>%
    filter(`Age Group` != "< 10")
  # Calculations
  total_screened <- sum(inc$n[inc$ogtt == 1])
  total_seen <- sum(inc$n)
  screen_rate <- total_screened / total_seen * 100

  screened_10_18 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "10 - 18"]
  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
  rate_10_18 <- screened_10_18 / seen_10_18 * 100

  screened_19_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "19 - 30"]
  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
  rate_19_30 <- screened_19_30 / seen_19_30 * 100

  screened_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "> 30"]
  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
  rate_30 <- screened_30 / seen_30 * 100

  # Nice results
  list(
    `Year` = year,
    `Total Screened` = total_screened,
    `Total Seen` = total_seen,
    `Screen Rate (%)` = screen_rate,
    `Total Screened Age 10-18` = screened_10_18,
    `Total Seen Age 10-18` = seen_10_18,
    `Screen Rate Age 10-18 (%)` = rate_10_18,
    `Total Screened Age 19-30` = screened_19_30,
    `Total Seen Age 19-30` = seen_19_30,
    `Screen Rate Age 19-30 (%)` = rate_19_30,
    `Total Screened Age > 30` = screened_30,
    `Total Seen Age > 30` = seen_30,
    `Screen Rate Age > 30 (%)` = rate_30
  )
}
screen <- lapply(2003:2018, function(y) {
  year_screen(year = y)
})
screen <- data.frame(do.call(rbind, screen), check.names = F)
screen <- data.frame(lapply(screen, as.numeric), check.names = F)
kable(screen, digits = 3)
```

```{r}
#| label: tbl-screening_a1c
#| tbl-cap: "Yearly HbA1c Screening Rates"
year_screen <- function(df = encounter, year) {
  # Filter and count
  inc <- df %>%
    filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
    mutate(a1c = !is.na(hgba1c)) %>%
    group_by(eDWID) %>%
    summarise(a1c = max(a1c), `Age Group` = `Age Group`[1])
  inc <- inc %>%
    group_by(a1c) %>%
    count(`Age Group`) %>%
    filter(`Age Group` != "< 10")
  # Calculations
  total_screened <- sum(inc$n[inc$a1c == 1])
  total_seen <- sum(inc$n)
  screen_rate <- total_screened / total_seen * 100

  screened_10_18 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "10 - 18"]
  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
  rate_10_18 <- screened_10_18 / seen_10_18 * 100

  screened_19_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "19 - 30"]
  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
  rate_19_30 <- screened_19_30 / seen_19_30 * 100

  screened_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "> 30"]
  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
  rate_30 <- screened_30 / seen_30 * 100

  # Nice results
  list(
    `Year` = year,
    `Total Screened` = total_screened,
    `Total Seen` = total_seen,
    `Screen Rate (%)` = screen_rate,
    `Total Screened Age 10-18` = screened_10_18,
    `Total Seen Age 10-18` = seen_10_18,
    `Screen Rate Age 10-18 (%)` = rate_10_18,
    `Total Screened Age 19-30` = screened_19_30,
    `Total Seen Age 19-30` = seen_19_30,
    `Screen Rate Age 19-30 (%)` = rate_19_30,
    `Total Screened Age > 30` = screened_30,
    `Total Seen Age > 30` = seen_30,
    `Screen Rate Age > 30 (%)` = rate_30
  )
}
screen <- lapply(2003:2018, function(y) {
  year_screen(year = y)
})
screen <- data.frame(do.call(rbind, screen), check.names = F)
screen <- data.frame(lapply(screen, as.numeric), check.names = F)
kable(screen, digits = 3)
```

In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by registry definition) and those under age 10 were excluded. Screening rate is calculated as the number of people with an OGTT or HbA1c, respectively, divided by the total number of people seen in clinic that year.

@@

```{r}
# Import data and codebook
annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
data_dict <- readxl::read_excel("./Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
# Remove those with T1D or T2D
annual <- annual %>% arrange(eDWID, ReviewYear)
annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
  levels = 1:3,
  labels = c("CFRD-", "CFRD-", "CFRD+")
)
annual$age_2018 <- 2018 - (annual$ReviewYear - annual$Age_YrEnd)
annual <- annual %>%
  group_by(eDWID) %>%
  mutate(age_2018 = as.numeric(names(sort(table(age_2018),
    decreasing = TRUE
  )[1])))
# Get mutations
demogs <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
demogs <- demogs %>% select(eDWID, Gender, MutClass)
annual <- left_join(annual, demogs, by = join_by(eDWID))
# Variable formatting
# If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
# on any supplemental feeding
# Include all kinds of tube feeds
annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
annual$tube_feeding <-
  apply(annual[, tube_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Liver disease
liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
  apply(annual[, liver_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Calculate percentiles
annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$age_perc <- ifelse(annual$Age_YrEnd < 20, annual$Age_YrEnd, 20)
annual$height_perc <- sds(annual$A_height,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "height2_20", ref = cdc.ref, type = "perc"
) * 100
annual$bmi_perc <- sds(annual$bmi,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "bmi", ref = cdc.ref, type = "perc"
) * 100
# Format columns
annual$MutClass <- factor(annual$MutClass,
  levels = c("1-3", "4-5", "Oth", "N/A"),
  labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
  labels = c("0", "1", "2", ">=3"),
  breaks = c(0, 1, 2, 3, Inf), right = F
)
annual$A_FEV1_pct_predicted <- cut(annual$A_FEV1_pct_predicted,
  breaks = c(0, 40, 80, 100, Inf), right = F
)
annual$A_FEV1_pct_predicted <- relevel(annual$A_FEV1_pct_predicted,
  ref = "[100,Inf)"
)
annual$height_perc <- cut(annual$height_perc, breaks = c(0, 10, 25, 50, Inf))
annual$height_perc <- relevel(annual$height_perc, ref = "(50,Inf]")
annual$bmi_perc <- cut(annual$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
annual$bmi_perc <- relevel(annual$bmi_perc, ref = "(50,85]")
annual$age_group <- cut(annual$Age_YrEnd,
  breaks = c(-Inf, 10, 18, 30, Inf),
  right = F
)
annual$age_group <- relevel(annual$age_group, ref = "[10,18)")
# Bugs
bugs <- c(
  "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_pulmonarycomplications1", "A_mycocultureresults3"
)
# Yes/No variables
yn_vars <- c(
  "A_IsOnEnzymes", "A_corticosteroids1", "A_feeding2",
  "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"
)
annual[, yn_vars] <- lapply(annual[, yn_vars], factor,
  levels = 0:1,
  labels = c("No", "Yes")
)
annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
# To have CFRD, must have two consecutive visits with CFRD
# Once diagnosed, always in CFRD group.
annual = annual %>%
  group_by(eDWID) %>%
  mutate(two_consec = 0)
```

@@

# Survival analysis

Based on CFFPR annualized data.

```{r}
# Create survival dataset
s <- annual %>%
  rename(end = ReviewYear) %>% # ReviewYear is the end of the interval
  group_by(eDWID) %>%
  mutate(start = lag(end,default = 2002)) %>% # Get interval start
  select(eDWID, start, end, Age_YrEnd, cfrd_status_annual) # Sort columns
# Age group
s$age_group <- cut(s$Age_YrEnd,
  breaks = c(-Inf, 10, 19, 31, Inf), right = F
)
knitr::kable(head(s,40))
```

```{r}
# New dataset approach
cases = unique(annual$eDWID[annual$cfrd_status_annual == "CFRD+"])
controls_df = annual[!annual$eDWID %in% cases,]
cases_df = annual[annual$eDWID %in% cases,]
# Controls
tc = controls_df %>% 
  select(eDWID,ReviewYear,cfrd_status_annual) %>%
  filter(!is.na(cfrd_status_annual)) %>%
  group_by(eDWID) %>%
  summarise(start = first(ReviewYear),
         end = last(ReviewYear),.groups = "drop") %>%
  mutate(cfrd = 0)
# Cases
tc2 = cases_df %>% 
  select(eDWID,ReviewYear,cfrd_status_annual) %>%
  filter(!is.na(cfrd_status_annual)) %>%
  mutate(neg_year = ReviewYear * (ifelse(cfrd_status_annual=="CFRD-",-1,1))) %>%
  arrange(eDWID,cfrd_status_annual,neg_year) %>%
  group_by(eDWID,cfrd_status_annual) %>% 
  filter(row_number()==1)
```

```{r}
# Survival fit
surv_obj <- Surv(
  time = s$start, time2 = s$end,
  event = as.numeric(s$cfrd_status_annual) - 1
)
fit <- survfit(surv_obj ~ age_group, id = eDWID, data = s)
ggsurvplot(fit,
  risk.table = T, cumevents = T,
  fun = "event", xlim = c(2002, 2018)
)
```
